Accurate binding of calcium to phospholipid bilayers by effective inclusion of electronic polarization

### 2017-10-10 Manuscript/manuscript.tex
New version of conclusions

\section{Conclusions}
By using the electrometer concept we show that the binding of \ce{Na^+} and \ce{Ca^2+} ions
to a POPC lipid bilayer can be accurately described with classical MD simulation 
force fields, where electronic polarization is implicitly included using 
the electronic continuum correction (ECC)~\cite{leontyev11}.
The proposed ECC-lipids model is a significant improvement over 
other available lipid models, which all overestimate specific cation binding affinities~\cite{catte16}.  
While the structural details of a POPC lipid bilayer simulated with the ECC-lipids
model agree with experiments with the comparable accuracy to the other state of the art lipid models,
it also reproduces the experimental lipid head group order parameter responses to
the cationic surfactant, NaCl and CaCl$_2$ concentrations.

The good agreement with experiments enables the atomic resolution 
interpretation of NMR experiments by using MD simulations.
In agreement with previous interpretations of experimental data \cite{hauser76,hauser78,herbette84,binder02},
the Ca$^{2+}$ ions mainly interact with phosphate oxygens.
However, the stoichiometry of the binding is significantly more complicated 
that the simple ternary complex model.
Direct analysis of the details of calcium binding from MD simulations is in agreement
with ternary complex model, which is suggested based on NMR data \cite{altenbach84}.
In this model one calcium binds to two POPC molecules, which together form a ternary
complex.
\todo{Continue summary using previous section once it is finished.}

The electronic continuum correction is built here upon the Lipid14 POPC model \cite{dickson14},
but we expect that the correction can be used also for other lipids
and force fields.
\todo{JOE: Do we want to suggest that ECC-lipids can be used 
with current force fields to study for example membrane proteins? 
This was expressed with the following sentence: 
The parameters can be used with existing standard nucleic acid and protein force fields, e.g. AMBER-FB15~\cite{Wang2017}. }
We suggest using state of the art water models like OPC3\cite{Izadi16} or OPC \cite{Izadi14},
which yield higher accuracy than the traditional TIP3p water model~\cite{jorgensen83}.
Several water models 
(OPC3\cite{Izadi16}, OPC \cite{Izadi14}, SPC/E \cite{Berendsen1987} and TIP4p/2005 \cite{Abascal2005}) 
were used to exemplify the transferability of 
the parameters of the new ECC-lipids force field.

@@

\newpage
\appendix

>>>>

\section{Conclusions}
By using the electrometer concept we show that the binding of \ce{Na^+} and \ce{Ca^2+} ions
to a POPC lipid bilayer can be accurately described with the classical MD simulation 
force field, where the electronic polarization is implicitly included using 
the electronic continuum correction (ECC)~\cite{leontyev11}.
The proposed ECC-lipids model is a significant improvement over 
other available lipid models, which all overestimate specific cation binding affinities~\cite{catte16}.  
While the structural details of a POPC lipid bilayer simulated with the ECC-lipids
model agree with experiments with the comparable accuracy to the other state of the art lipid models,
it also reproduces the experimental lipid head group order parameter responses to
the cationic surfactant, NaCl and CaCl$_2$ concentrations.

The good agreement with experiments enables the atomic resolution 
interpretation of NMR experiments by using MD simulations.
In agreement with previous interpretations of experimental data \cite{hauser76,hauser78,herbette84,binder02},
the Ca$^{2+}$ ions mainly interact with phosphate oxygens.
However, the stoichiometry of the binding is significantly more complicated 
than in the previous interpretation of the NMR data based on 
the ternary complex model, where one calcium binds to two POPC molecules~\cite{altenbach84}.
The complexes of one calcium bound to two lipids are the most probable also in the
ECC-lipids model, but the complexes of one or three lipids per one calcium
were observed to be almost equally likely. While the success of the ternary
compex model is understandable based on the simulation results,
a simple binding model cannot detec the complex binding
oberved in the simulation.

The improved cation binding behaviour to POPC bilayer pave the way for
simulations of complex biochemical systems with correctly described
electrostatic interactions in the vicinity of cellular membranes.
The ECC-lipids model is build by scaling the partial charges and
the LJ-radius of the headgroup, glycerol backbone and carbonyl atoms of
the Lipid14 POPC model~\cite{dickson14}. While the Lipid14 model
is compatible with the AMBER force field family, the compatibility of
the ECC-lipids model may be compromised due to the changes in intermolecular
interactions of the scaled atoms. On the other hand, a fully consistent
ECC-force field should include the correction also in other than lipid
molecules, including water. The work toward this direction and the extension
to other lipid molecules and force fields is left for the future studies.

%but we expect that the correction can be used also for other lipids
%and force fields.

%We suggest using state of the art water models like OPC3\cite{Izadi16} or OPC \cite{Izadi14},
%which yield higher accuracy than the traditional TIP3p water model~\cite{jorgensen83}.
%Several water models 
%(OPC3\cite{Izadi16}, OPC \cite{Izadi14}, SPC/E \cite{Berendsen1987} and TIP4p/2005 \cite{Abascal2005}) 
%were used to exemplify the transferability of 
%the parameters of the new ECC-lipids force field.

@@

\newpage
\newpage
\appendix
